---
title: "Untitled"
output:
  pdf_document: default
  html_document: default
date: "2024-04-03"
---

```{r, echo=FALSE, message=FALSE}
suppressWarnings(suppressMessages({
  # install.packages("knitr") # install the knitr package
  library(knitr) # load the knitr package
  
  library(dplyr)
  library(tidyr)
  library(caret)
  # 'gridExtra': arrangee the plots in 2x2 layout 
  
  # install.packages('gridExtra')
  library(gridExtra)
  
  # install.packages('corrplot')
  # library(corrplot)
  # remove.packages("ggplot2") # Unisntall ggplot
  
  # install.packages("ggplot2") # Install it again
  library(ggplot2) # Load the librarie (you have to do this one on each new session)
  
  # install.packages("glmnet")
  library(glmnet)
  
  # install.packages('car')
  library(car)
  
  library(pROC)
  
  # install.packages('faraway')
  library(faraway)
}))
```

```{r}
set.seed(1007217101)
```


## Load Data

```{r,echo=F}

dt0 <- read.csv("toronto.csv", header = TRUE)

# 1. Scan 'date_time' column, drop data from year 2008
dt0 = dt0 %>% filter(!grepl("2008", date_time))
# dt0 <- subset(dt0, !grepl("2008", dt0$date_time))


## Generate Y variable 'rained'
dt0$precipMM <- as.numeric(dt0$precipMM) # justincase 'precipMM' is not formatted as numeric
dt0$rained <- ifelse(dt0$precipMM > 0, 1, 0)


## Generate cat. variable 'snowed'
dt0$totalSnow_cm <- as.numeric(dt0$totalSnow_cm) # justincase not numeric
dt0$snowed <- ifelse(dt0$totalSnow_cm > 0, 1, 0)


## Generate  cat. variable 'season' --> further, generate indicator whether it is winter

# Ensure correct Date format
dt0$date_time <- as.Date(dt0$date_time)
# convert  to month number
dt0$month <- as.numeric(format(dt0$date_time, "%m"))
# assign seasons based on the month
dt0$season <- cut(dt0$month, breaks=c(0, 3, 5, 8, 11, 12), labels=c("winter", "spring", "summer", "fall", "winter"), include.lowest=TRUE)
# Createbinary variable 'is_winter'
dt0$is_winter <- ifelse(dt0$season == 'winter', 1, 0)



## Generate 'medTempC'
dt0$medTempC <- apply(dt0[,c('maxtempC', 'mintempC')], 1, median)


dt0 <- na.omit(dt0)
```

```{r,echo=F,include=F}
write.csv(dt0,"Toronto_clean.csv")
```

```{r}
dt <- read.csv("Toronto_clean.csv", header = TRUE)
dt <- subset(dt, select = -X)
dt$rained <- as.factor(dt$rained)
dt$snowed <- as.factor(dt $snowed)
dt$is_winter <- as.factor(dt $is_winter)

dt %>% glimpse()
dt %>% colnames()
dt %>% head(10) # 4048 rows * 31 cols
```

## Mutates

```{r}
dt <- dt %>% mutate(log_visibility = log(visibilityKM))
# p<-ggplot(as_tibble(dt), aes(x=log_visibility, fill=rained)) +
#       geom_histogram(position="dodge", binwidth=1) +
#       labs(x="Visibility (km)", y="Count of Rainy Days")
# p
```


## Split Data Train Test 8:2.  But this step is not necessary in this study as we will be using lrm() to validate, which automatically conducts k-fold cross-validation. At this stage, we would just define variable spaces.


```{r}
# train_indexes <- sample(1:nrow(dt), size = 0.8 * nrow(dt)) # 80% for training
# dt_train <- dt[train_indexes, ]
# dt_test <- dt[-train_indexes, ]
# 
# # dt_train %>% glimpse() #3238
# # dt_test %>% glimpse() #810



full_predictors <- c('medTempC' , "humidity",
                      'cloudcover', 'windspeedKmph' , 'log_visibility' , 'pressure' , 
                      'DewPointC' , 'sunHour' , 'uvIndex' , 'WindGustKmph' ,
                      'winddirDegree' ,'moon_illumination_percent' ,'FeelsLikeC')
# full_predictors <- c(full_predictors,'rained')
full_predictors
# dt['rained'] is the response variable.

```

## Correlation Matrix to first select candidate featrues (Not included)

```{r}
# numer = c('medTempC' , "humidity" , 
#                       'cloudcover', 'windspeedKmph' , 'visibilityKM' , 'pressure' , 
#                       'DewPointC' , 'sunHour' , 'uvIndex' , 'WindGustKmph' ,
#                       'winddirDegree' ,'moon_illumination_percent' ,'FeelsLikeC')
# 
# cmatrix <- cor(dt[numer])
# dt[numer]
# #col <- colorRampPalette(c())
# corrplot(cmatrix,addCoef.col="grey",number.cex=0.5,tl.cex=0.6)
```

## Full model)

```{r}
model1 <- glm(rained ~ .,
              family = binomial(link = logit), 
              data = dt[c(full_predictors,'rained')])

model1 %>% summary()
```


## Perform AIC based stepwise selection 

```{r, eval=TRUE, echo = T}
## Stepwise elimination based on AIC ##
sel.var.aic <- step(model1, trace = 0, k = 2, direction = "both") 
select_var_aic<-attr(terms(sel.var.aic), "term.labels")   
select_var_aic
```


## Now perform BIC based selection

```{r, eval=TRUE, echo = T}
## Stepwise elimination based on AIC ##
sel.var.bic <- step(model1, trace = 0, k = log(nrow(dt)), direction = "both") 
select_var_bic<-attr(terms(sel.var.bic), "term.labels")   
select_var_bic
```


## Lasso Selection

Similar process: first we fit model with differnt lambdas

```{r}
X <- as.matrix(dt[full_predictors])
Y <-dt$rained

# grid = 10^seq(10,-2,length = 100)

cv.out <- cv.glmnet(X,Y,alpha=1, family= 'binomial') # 10 fold cross validation
bestlam <- cv.out$lambda.min
bestlam


#Then we look at the best model i.e. who has the least lambda. And extract its variables.


lasso.mod <- glmnet(X,Y,alpha=1,lambda=bestlam,family= 'binomial') #get the model under the best lambda value
coefs <-coef(lasso.mod)[,1]
coefs<-coefs[coefs != 0]
a <-coefs %>% as.data.frame()
select_var_lasso =a %>% rownames()
select_var_lasso=select_var_lasso[!select_var_lasso %in% c("(Intercept)")] 
select_var_lasso
```

### A Helper Function for df beta:

```{r}

Draw_dfbetas <- function(Features, y='rained'){
  modeltemp <- glm(rained ~ .,
              family = binomial(link = logit), data = dt[Features]) 
  # modeltemp %>% summary
  df.final <- dfbetas(modeltemp)
  
  for (feature in Features[!Features %in% 'rained']) {
    
    # df.final %>% head
    
    par(family = 'serif')
    plot(dt[,feature], df.final[,feature], xlab=feature, 
         ylab='dfbeta')
    lines(lowess(dt[,feature], df.final[,feature] ), lwd=2, col='blue')
    abline(h=0, lty='dotted')
    abline(h=-2/sqrt(nrow(df.final)), lty='dotted')
    abline(h=2/sqrt(nrow(df.final)), lty='dotted')
  }
}

# Draw_dfbetas(Features = candidates_6)
```


### Helper Function for Deviance residuals
```{r, eval=TRUE, echo = T}

# ## Plot the deviance residuals ##
# res.dev <- residuals(model2, type = "deviance")
# par(family = 'serif')
# plot(dt[,'WindGustKmph'], res.dev, xlab='WindGustKmph', 
#      ylab='Deviance Residuals')
# lines(lowess(dt[,'WindGustKmph'], res.dev), lwd=2, col='blue')
# abline(h=0, lty='dotted')
# 



Draw_devianceResidual <- function(Features, y='rained'){
  modeltemp <- glm(rained ~ .,
              family = binomial(link = logit), data = dt[Features]) 
  # modeltemp %>% summary
  df.final <- dfbetas(modeltemp)
  
  for (feature in Features[!Features %in% 'rained']) {
    
    res.dev <- residuals(modeltemp, type = "deviance")
    par(family = 'serif')
    plot(dt[,feature], res.dev, xlab=feature, 
         ylab='Deviance Residuals')
    lines(lowess(dt[,feature], res.dev), lwd=2, col='blue')
    abline(h=0, lty='dotted')
  }
}
# Draw_devianceResidual(candidates_6)
```






## Define Variable Space for the final model


```{r}
candidates_ab <- intersect(select_var_aic, select_var_bic)
candidates_abl <- intersect(candidates_ab, select_var_lasso)
candidates_bl <- intersect(select_var_bic, select_var_lasso)

candidates_1 <- candidates_abl[!candidates_abl %in% c('uvIndex','WindGustKmph')]
candidates_1 = c(candidates_1,'humidity' )

candidates_2 = candidates_1 = c(candidates_1,'is_winter','snowed' )
candidates_2 <- candidates_2[!candidates_2 %in% c('windspeedKmph','snowed')]
# dt$windspeedKmph
# candidates_2 <- sel.var.b2[!sel.var.b2 %in% c('WindGustKmph')]
# candidates_2 = c(candidates_2,'humidity' )

# 'WindGustKmph','uvIndex'
# # candidates_6 = c(candidates_4,'log_visibility' )
# # candidates_6= candidates_6[!candidates_6 %in% c("visibilityKM")]
# 
# candidates_7 = c(full_predictors,'rained','log_visibility')
# candidates_7 = candidates_7[!candidates_7 %in% c('FeelsLikeC',"DewPointC",'moon_illumination_percent','WindGustKmph','visibilityKM')]

```



## Plot the dfbetas and deviance residuals


### Fit the initial final model, and plot deviance betas 

```{r}

ft = candidates_2
ftr = c(ft,'rained')

modelF <- glm(rained ~ .,family = binomial(link = logit), data = dt[ftr])
# Draw_dfbetas(Features = ftr)
# Draw_devianceResidual(ftr)

modelF %>% summary()
modelF%>% vif()
```



## Check and remove outliers

```{r}


# Extract linear predictors (eta)
eta <- predict(modelF, type = "link")

# Calculate standardized residuals
residuals_standardized <- rstandard(modelF)

# Create QQ plot
qqnorm(residuals_standardized)
qqline(residuals_standardized, col = "red")

######
plot(eta, resid(modelF, type = "deviance"),
     xlab = "Linear Predictor", ylab = "Deviance Residuals", main = "")

# Deviance Residuals vs. Fitted Values
plot(fitted(modelF), resid(modelF, type = "deviance"),
     xlab = "Fitted Values", ylab = "Deviance Residuals", main = "")

#identify potential outliers with absolute standardized residuals greater than 2
potential_outliers <- which(abs(residuals_standardized) > 2)

#Find  potential outliers
outlier_data <- modelF$data[potential_outliers, ]
outlier_data

#Remove them:
dt2 <- modelF$data[-potential_outliers, ]

```

## Fit again, as the final Model:

```{r}
ft = candidates_2
ftr = c(ft,'rained')
ftr
modelFF <- glm(rained ~ .,family = binomial(link = logit), data = dt2[ftr])
modelFF %>% summary()
modelFF%>% vif()

```



## Model calibration with cross-validation and bootstrap. 

### Plus the QQ-Plot and Deviance Residual plot, which will be cobined to display in the report.

```{r, eval=TRUE, echo = T,fig.height=8,fig.width=9 }
library(rms)

par(mfrow = c(2,2))

## Fit the model with lrm from rms package ##
lrm.final <- lrm(rained ~ ., data = dt2[ftr], x =TRUE, y = TRUE, model= T)
# cross.calib <- calibrate(lrm.final, method="crossvalidation", B=10) # model calibration
cross.calib <- calibrate(lrm.final, method="crossvalidation", B=10) # model calibration
plot(cross.calib, las=1, xlab = "Predicted Probability")

## Discrimination with ROC curve 

# library(pROC)
p <- predict(lrm.final, type = "fitted")

roc_logit <- roc(dt2$rained ~ p)
## The True Positive Rate ##
TPR <- roc_logit$sensitivities
## The False Positive Rate ##
FPR <- 1 - roc_logit$specificities

plot(FPR, TPR, xlim = c(0,1), ylim = c(0,1), type = 'l', lty = 1, lwd = 2,col = 'red')
abline(a = 0, b = 1, lty = 2, col = 'blue')
text(cex = 1.2,0.5,0.5,label = paste("AUC = ", round(auc(roc_logit),2)))

auc(roc_logit)
qqnorm(residuals_standardized)
qqline(residuals_standardized, col = "red")

# Deviance Residuals vs. Fitted Values
plot(fitted(modelF), resid(modelF, type = "deviance"),
     xlab = "Fitted Values", ylab = "Deviance Residuals", main = "")
```


## Generate summary table for final data. Exported and pasted into report document.


```{r}
# install.packages("psych") 
library(psych)


# write.csv(describe(dt2), file = "summary_table.csv", row.names = T)

mean(dt2$is_winter==1)
# modelFF$coefficients
```
## Summary table for all candidate models

```{r}
suppressWarnings(suppressMessages({
  modelA<-glm(rained ~ .,family = binomial(link = logit), data = dt[c(select_var_aic,'rained')])
  modelB<-glm(rained ~ .,family = binomial(link = logit), data = dt[c(select_var_bic,'rained')])
  modelL<-glm(rained ~ .,family = binomial(link = logit), data = dt[c(select_var_lasso,'rained')])
  
  # modelsn <- glm(rained ~ .,family = binomial(link = logit), data = dt[c(candidates_2,'rained')])
  
  # install.packages('stargazer')
  library(stargazer)
  
  stargazer(modelFF, modelA, modelB, modelL, type='text', digits = 4, title = 'Table 2: Summary of All Candidate Models', style = 'qje', column.labels = c('Final Model', 'AIC Selected', 'BIC Selected', 'Lasso Selected'), dep.var.labels = 'Models')
}))
```

## The following are EDA plots, which will be displayed in appendix of the report.

```{r}
suppressWarnings(suppressMessages({

p1 <- ggplot(as_tibble(dt2), aes(x=medTempC, fill=rained)) +
      geom_histogram(position="dodge", binwidth=2) +
      labs(x="Median Temperature (C)", y="Count of Rainy Days")

p2 <- ggplot(as_tibble(dt2), aes(x=humidity, fill=rained)) +
      geom_histogram(position="dodge", binwidth=2) +
      labs(x="Humidity Level (%)", y="Count of Rainy Days")

p3 <- ggplot(as_tibble(dt2), aes(x=pressure, fill=rained)) +
      geom_histogram(position="dodge", binwidth=2) +
      labs(x="Pressure", y="Count of Rainy Days")

p4 <- ggplot(as_tibble(dt2), aes(x=log_visibility, fill=rained)) +
      geom_histogram(position="dodge", binwidth=2) +
      labs(x="Logged Visibility", y="Count of Rainy Days")

p5 <- ggplot(as_tibble(dt2), aes(x=cloudcover , fill=rained)) +
      geom_histogram(position="dodge", binwidth=1) +
      labs(x="Cloud coverage rate", y="Count of Rainy Days")

bar1 <- ggplot(dt2, aes(x=rained, fill=is_winter)) +
    geom_bar(position="dodge", binwidth=5)+ 
    labs(x="Rained", y="Count of Days", fill="Winter")

grid.arrange(p1, p2, p3, p4, p5, bar1, nrow=2)
}))
```
### Boxplots for numeric variables
```{r, echo=FALSE}
suppressWarnings(suppressMessages({

  pb1 <- ggplot(as_tibble(dt2), aes(x=rained, y=medTempC)) +
        geom_boxplot() +
        labs(x="Rain Occurrence", y="Median Temperature (C)")
  
  pb2 <- ggplot(as_tibble(dt2), aes(x=rained, y=humidity)) +
        geom_boxplot() +
        labs(x="Rain Occurrence", y="Humidity Level (%)")
  
  pb3 <- ggplot(as_tibble(dt2), aes(x=rained, y=pressure)) +
        geom_boxplot() +
        labs(x="Rain Occurrence", y="Pressure (mb)")

  pb4 <- ggplot(as_tibble(dt2), aes(x=rained, y=log_visibility)) +
        geom_boxplot() +
        labs(x="Rain Occurrence", y="Logged Visibility (Km)")  
  pb5 <- ggplot(as_tibble(dt2), aes(x=rained, y=cloudcover)) +
        geom_boxplot() +
        labs(x="Rain Occurrence", y="Cloud Cover (%)")
  

  
  
  grid.arrange(pb1, pb2, pb3, pb4, pb5, nrow=2)

}))
```
```{r}
# grid.arrange(p1, p2, p3, p4, p5, bar1, 
#              pb1, pb2, pb3, pb4, pb5, nrow=2)
```

